---
title: "Covid Dataset"
author: "Justin Nguyen"
date: "4/10/2021"
output: html_document
---

Packages
```{r} 
install.packages("tidyverse", "zoo", "readxl","corrplot")
install.packages("rpart")
install.packages("MASS")
install.packages("cluster")
install.packages("caTools")
install.packages("randomForest")
install.packages("rpart.plot")
```

Library
``` {r}
library(tidyverse)
library(rpart)
library(readxl)
library(corrplot)
library(MASS)
library(zoo)
library(cluster)
library(caTools)
library(randomForest)
library(rpart.plot)
```

Install File
``` {r}
covid_db_raw <- as.data.frame(read.csv("covid19.csv"))
healthcare_db_raw <- as.data.frame(read_excel("Country Healthcare Performance.xlsx"))
```

Data Cleaning

```{r}


covid_db_vax <- covid_db_raw %>% 
  mutate(Sinovac = case_when(str_detect(vaccines, "Sinovac") ~ 1,
                             str_detect(vaccines, "Sinopharm") ~ 1),
         Sputnik = case_when(str_detect(vaccines, "Sputnik") ~ 1),
         Johnson = case_when(str_detect(vaccines, "Johnson&Johnson")~ 1),
         Pfizer = case_when(str_detect(vaccines, "Pfizer") ~ 1),
         Moderna = case_when(str_detect(vaccines, "Moderna") ~ 1),
         AstraZeneca = case_when(str_detect(vaccines, "Oxford") ~ 1),
         Covaxin = case_when(str_detect(vaccines, "Covaxin") ~ 1)) %>%
  dplyr::select(country, Sinovac, Sputnik, Johnson, Pfizer, Moderna, AstraZeneca, Covaxin) %>%
  replace(is.na(.), 0) %>%
  mutate(Total_Vax_Type = Sinovac  + Sputnik + Johnson + Pfizer + Moderna + AstraZeneca + Covaxin) 

covid_db_full_clean <- covid_db_raw %>%
  left_join(covid_db_vax, by = "country") %>% 
  mutate(GDP_Bracket = ifelse(GDP_per_capita >= 40000, "High GDP Per Cap", 
                              ifelse(GDP_per_capita >= 30000, "Upper Middle GDP Per Cap",
                                     ifelse(GDP_per_capita >= 20000, "Middle GDP Per Cap",
                                            ifelse(GDP_per_capita >= 10000, "Lower Middle GDP Per Cap", "Low GDP Per Cap"))))) %>%
  left_join(healthcare_db_raw, by = c( "country" = "Country"))

```
To dO: 

  1. Develop and Test Normalized Model outcomes such as people vaccinated per hundred (Done)
  2. Connect Findings to all analysis (Done)
  3. Test correlations between variables for interaction and for multi collinearity (Done)
  4. Explore other Tree
  5. Consider whether to use regression trees or cluster modeling
  6. Add lessons learned and findings 
  7. Research additional modeling 
  
Descriptive Analytics 
``` {r}
covid_db_full_clean %>%
  ggplot(aes(x = Health_expenditure, y = total_vaccinations_per_hundred, col = GDP_Bracket)) +
  geom_point() + 
  scale_y_continuous(trans = 'log10')


#Descriptive Analysis points for Presentation
  #Find by Bracket, Make note of why we did use region

  #Population, GDP Per Cap, Health Expenditure, Total Vaccinations, People Vaccinations, Total Vax Type

covid_db_full_clean %>%
  group_by(GDP_Bracket) %>%
  filter(!is.na(GDP_Bracket)) %>%
  summarise(Avg_Pop = mean(population, na.rsdm = T),
            SD_Pop = sd(population, na.rm = T),
            Avg_GDP_Per_Cap = mean(GDP_per_capita ,na.rm = T),
            SD_GDP_Per_Cap = sd(GDP_per_capita ,na.rm = T),
            AVG_Health_Expenduture = mean(Health_expenditure, na.rm = T),
            SD_Health_Expenduture = sd(Health_expenditure, na.rm = T)
            ,
            AVG_Total_Vaccination = mean(total_vaccinations),
            AVG_people_Vaccinated = mean(people_vaccinated, na.rm = T),
            AVG_Vaccine_Utilization = mean(Total_Vax_Type))
#Note Max Min countries and show notable countries 

covid_db_full_clean %>%
  filter(country %in% c("United States of America", "United Kingdom", "China", "Russia", "Japan", "South Africa", "Israel", "Germany", "India", "Brazil", "Grenada")) %>%
  group_by(country) %>%
  summarise(Avg_Pop = mean(population, na.rm = T),
            Avg_GDP_Per_Cap = mean(GDP_per_capita ,na.rm = T),
            AVG_Health_Expenduture = mean(Health_expenditure, na.rm = T),
            AVG_Total_Vaccination = mean(total_vaccinations),
            AVG_people_Vaccinated = mean(people_vaccinated, na.rm = T),
            AVG_Vaccine_Utilization = mean(Total_Vax_Type))

chisq.test(table(covid_db_full_clean$GDP_Bracket))
chisq.test(table(covid_db_full_clean$GDP_Bracket), table(covid_db_full_clean$Healthcare_System))

t.test(GDP_per_capita ~ Pfizer, data = covid_db_full_clean)

anova(aov(people_vaccinated_per_hundred ~ GDP_growth_rate * CO2_emission, data = covid_db_full_clean))

```


visual Analytics 
```{r}

covid_db_full_clean %>%
  ggplot(aes(x = population, y = total_vaccinations_per_hundred, col = region)) +
  geom_point() +
  scale_x_continuous(trans = 'log10') + 
  scale_y_continuous(trans = 'log10')+
  geom_abline()
#Should consolidate the Regions 

#Histogram Stopgap Method
x <- covid_db_full_clean %>% 
  summarise(Sinovac = sum(Sinovac),
            Sputnik = sum(Sputnik),
            Johnson = sum(Johnson),
            Pfizer = sum(Pfizer),
            Moderna = sum(Moderna),
            Oxford = sum(AstraZeneca),
            Covaxin = sum(Covaxin)) %>%
  pivot_longer(everything()) 

barplot(x$value, 
        xlab = "Vaccine", 
        ylab = "# of Countries",
        main = "Number of Applications of Covid 19 Vaccines",
        names.arg = x$name, 
        col = "dark blue")

covid_db_full_clean %>% #Need to determine What the heck is employment Industry 
  ggplot(aes(x = employment_industry, y = total_vaccinations, col = GDP_Bracket)) +
  geom_point() +
  scale_y_log10() 


covid_db_full_clean %>%
  filter(!is.na(GDP_Bracket) & !is.na(people_vaccinated_per_hundred)) %>%
  ggplot(aes(x = GDP_Bracket, y = people_vaccinated_per_hundred)) +
  geom_boxplot() +
  scale_y_continuous(limits = c(0, 40)) + 
  coord_flip() +
  ggtitle("Boxplot: GDP_Bracket Versus People Vaccinated per Hundreds", "Outliers over 40 Removed") + 
  theme_classic()

covid_db_full_clean %>%
  group_by(GDP_Bracket) %>%
  summarise(max(people_vaccinated_per_hundred, na.rm = T),
            mean(Total_Vax_Type))

covid_db_full_clean %>%
  filter(!is.na(GDP_Bracket)) %>%
  ggplot(aes(x = population, y = people_vaccinated_per_hundred, col = GDP_Bracket)) +
  geom_point() + 
  scale_x_log10() + 
  geom_abline() +
  xlab("Log(Population)") +
  ggtitle("Log of Population Versus People Vaccinated", "Countries with no GDP per Capita removed") +
  theme_classic()

covid_db_full_clean %>%
  filter(!is.na(GDP_Bracket)) %>%
  ggplot(aes(x = population, y = people_vaccinated_per_hundred, col = Healthcare_System)) +
  geom_point() + 
  scale_x_log10() + 
  geom_abline() +
  xlab("Log(Population)") +
  ggtitle("Log of Population Versus People Vaccinated", "Filtered by Healthcare Model Type") +
  theme_classic()

#Histogram

covid_db_full_clean %>% 
  filter(!is.na(people_vaccinated_per_hundred)) %>% 
  ggplot(aes(x = people_vaccinated_per_hundred)) + 
  geom_histogram(binwidth = 5, colour="black", fill="white") 

covid_db_full_clean %>% 
  filter(!is.na(people_vaccinated)) %>% 
  ggplot(aes(x = people_vaccinated)) + 
  geom_histogram(colour="black", 
                 fill="white") 

#Histogram with layers
covid_db_full_clean %>% 
  filter(!is.na(people_vaccinated_per_hundred)) %>% 
  ggplot(aes(x = people_vaccinated_per_hundred, fill = GDP_Bracket)) + 
  geom_histogram(binwidth = 15, 
                 position = "dodge",
                 alpha=.5) 

covid_db_full_clean %>% 
  filter(!is.na(people_vaccinated)) %>% 
  ggplot(aes(x = people_vaccinated, fill = GDP_Bracket)) + 
  geom_histogram( position = "identity",alpha=.5) 

```

Predictive Analytics 
```{r}
#Correlation Matrix
covid_db_train <- covid_db_full_clean %>% 
  drop_na()
corrplot(cor(covid_db_train[,c(4:15,17:20,22:23,25:29,31:32)]))
#Interesting correlations
  #Covaxin has strong positive correlation with population 
  #Pfizer and moderna has strong positive correlation with Strong GDP per cap
    #Pfizer has strong negative correlation with country with larger younger populations
  #Covaxin and Astrazeneca has strong correlation with total vax
  #People vaccinated has strong correlation with older pops, gdp per gap (weak), population,  CO2 Emmission, Health Expenditure, covaxin?, 

#Break Data down into testing data and training data


#Regression
summary(lm(people_vaccinated~ population + Population_age_60_and_above + CO2_emission + GDP_per_capita, data = covid_db_train))

summary(lm(people_vaccinated_per_hundred~ population + Population_age_60_and_above + CO2_emission , data = covid_db_train))
#Regression with interaction

best_model <- summary(lm(people_vaccinated ~ population*Population_age_60_and_above*GDP_per_capita + CO2_emission, data = covid_db_train))
  #Best result so far Adjusted R of .93
  #Shows strong correlation between how much of the population is over 60
  #Best Model
best_model

plot(people_vaccinated~ population, data = covid_db_full_clean, main = "Multivariate Interaction Model on People Vaccinated", xlab = "Population", ylab = "People Vaccinated", col = "Dark Blue")
abline(lm(people_vaccinated ~ population*Population_age_60_and_above*GDP_per_capita + CO2_emission, data = covid_db_train))

#prediction

covid_db_full_clean %>%
  mutate(people_vaccinated_predicted = -2.012*10^6 + 1.548*10^1*population + 1.149*10^5*Population_age_60_and_above + 1.239*10^2 * GDP_per_capita + -3.977*10^2*CO2_emission + -8.804*10^-1 *Population_age_60_and_above*population + 1.311*10^-2*population*GDP_per_capita + -5.877 *Population_age_60_and_above*GDP_per_capita + -3.608*10^-4*population*Population_age_60_and_above*GDP_per_capita,
         Difference = people_vaccinated - people_vaccinated_predicted,
         Ratio = people_vaccinated_predicted/(population*1000)) %>%
  filter(country %in% c("China", "Thailand", "Mexico")) %>%
  dplyr::select(country, population, Population_age_60_and_above, GDP_per_capita, CO2_emission, people_vaccinated_predicted, Ratio) %>%
  view()

covid_db_full_model<- covid_db_full_clean %>%
  mutate(people_vaccinated_predicted = -2.012*10^6 + 1.548*10^1*population + 1.149*10^5*Population_age_60_and_above + 1.239*10^2 * GDP_per_capita + -3.977*10^2*CO2_emission + -8.804*10^-1 *Population_age_60_and_above*population + 1.311*10^-2*population*GDP_per_capita + -5.877 *Population_age_60_and_above*GDP_per_capita + -3.608*10^-4*population*Population_age_60_and_above*GDP_per_capita,
         Difference = people_vaccinated - people_vaccinated_predicted,
         Ratio = people_vaccinated_predicted/(population*1000)) 

#New Model
summary(lm(people_vaccinated ~ population*GDP_per_capita + CO2_emission, data = covid_db_train))
best_model <- lm(people_vaccinated ~ population*GDP_per_capita + CO2_emission, data = covid_db_train)
  #By Removing People by 60 it removes a variable that was in a different form factor and was multicollinear 
summary(lm(people_vaccinated_per_hundred ~ population*GDP_per_capita + CO2_emission, data = covid_db_train))


covid_db_full_clean %>% 
  mutate(Predicted_People_Vaccinated = predict(best_model, newdata = covid_db_full_clean, type = "response"),
         Difference = Predicted_People_Vaccinated - people_vaccinated) %>%
  group_by(GDP_Bracket) %>% 
  summarise(mean(Predicted_People_Vaccinated, na.rm = T),
            mean(Difference, na.rm = T)) %>% 
  view()
summary(lm(people_vaccinated_per_hundred ~ GDP_per_capita + Urban_population + Total_Vax_Type, data = covid_db_train))

covid_db_full_clean %>% 
  mutate(Predicted_People_Vaccinated = predict(best_model, newdata = covid_db_full_clean, type = "response"),
         Difference = Predicted_People_Vaccinated - people_vaccinated) %>%
  filter(country %in% c("China", "Mexico", "Thailand")) %>%
  dplyr::select(country,
         population,
         GDP_per_capita,
         Predicted_People_Vaccinated)



```

Other Models
```{r}
#predicting Function 

vaccine_predictor <- function(vaccine_type, model) {
  Predicted_Decision <- predict(model, test_data, type = 'class')
  table(vaccine_type, Predicted_Decision)
}

#Kmeans
covid_vaccines <- covid_db_vax
covid_vaccines$Total_Vax_Type <- NULL

test <- kmeans(covid_vaccines[2:7], centers = 7)

clusplot(covid_vaccines, test$cluster, color = T, shade = T, lines = 0)

x <- covid_db_full_clean %>% 
  dplyr::select(GDP_per_capita, people_vaccinated, Urban_population, Total_Vax_Type, population) %>%
  drop_na()
test_2 <- kmeans(x, centers = 5)
clusplot(x, test_2$cluster, color = T, shade = T, lines = 0)


#Random Forest 
#Replace Random Forest with Trams' version 


#Training and Testing Data
sample <- sample.split(test_random$people_vaccinated_per_hundred, SplitRatio = .75)
train_data <- subset(test_random, sample == T)
test_data <- subset(test_random, sample == F)

randon_forest_1 <- randomForest(people_vaccinated_per_hundred ~. ,data = train_data)
prediction_1 <- predict(randon_forest_1, newdata = test_data[-17])
table(test_data[,17], prediction_1)



#Regression Tree for vaccine
Pfizer_Tree <- train_data %>% 
  dplyr::select(-total_vaccinations, -people_vaccinated, -total_vaccinations_per_hundred, -people_vaccinated_per_hundred, -Total_Vax_Type) %>%
  rpart(formula = Pfizer ~ ., 
                     method = "class")
rpart.plot(Pfizer_Tree, main = "Pfizer Adoption")
vaccine_predictor(test_data$Pfizer, Pfizer_Tree)

Moderna_Tree <- train_data %>% 
  dplyr::select(-total_vaccinations, -people_vaccinated, -total_vaccinations_per_hundred, -people_vaccinated_per_hundred, -Total_Vax_Type) %>%
  rpart(formula = Moderna ~ ., 
                     method = "class")
rpart.plot(Moderna_Tree, main = "Moderna Adoption")
vaccine_predictor(test_data$Moderna, Moderna_Tree)

Sinovac_Tree <- train_data %>% 
  dplyr::select(-total_vaccinations, -people_vaccinated, -total_vaccinations_per_hundred, -people_vaccinated_per_hundred, -Total_Vax_Type) %>%
  rpart(formula = Sinovac ~ ., 
                     method = "class")
rpart.plot(Sinovac_Tree, main = "Sinovac Adoption")

AstraZeneca_Tree <- train_data %>% 
  dplyr::select(-total_vaccinations, -people_vaccinated, -total_vaccinations_per_hundred, -people_vaccinated_per_hundred, -Total_Vax_Type) %>%
  rpart(formula = AstraZeneca ~ ., 
                     method = "class")
rpart.plot(AstraZeneca_Tree, main = "Astrazeneca Adoption")

Johnson_Tree <- train_data %>% 
  dplyr::select(-total_vaccinations, -people_vaccinated, -total_vaccinations_per_hundred, -people_vaccinated_per_hundred, -Total_Vax_Type) %>%
  rpart(formula = Johnson ~ ., 
                     method = "class")
rpart.plot(Johnson_Tree, main = "J&J Adoption")

Sputnik_Tree <- train_data %>% 
  dplyr::select(-total_vaccinations, -people_vaccinated, -total_vaccinations_per_hundred, -people_vaccinated_per_hundred, -Total_Vax_Type) %>%
  rpart(formula = Sputnik ~ ., 
                     method = "class")
rpart.plot(Sputnik_Tree, main = "Sputnik Adoption")

predict(AstraZeneca_Tree, test_data, type = 'class') 
table(test_data$AstraZeneca, predict(AstraZeneca_Tree, test_data, type = 'class'))

Covaxin_Tree <- train_data %>% 
  dplyr::select(-total_vaccinations, -people_vaccinated, -total_vaccinations_per_hundred, -people_vaccinated_per_hundred, -Total_Vax_Type) %>%
  rpart(formula = Covaxin ~ ., 
                     method = "class")
rpart.plot(Covaxin_Tree, main = "Covaxin Adoption")





```
